Zohmg---a large scale data store for aggregated time-series-based data. [DRAFT]

### 2009-08-12 doc/report/msc-report.tex
Report: abstract, preface, glossary, homeless text
Added swedish abstract, preface, and glossary words.
Removed Appendix C, homeless text.

\noindent {\Large \textbf{Abstract}}
\vspace{12pt}
\noindent

@@

Analyzing data at a massive scale is one of the biggest challenges that
Last.fm is facing. Interpreting patterns in user behaviour becomes a challenge
when millions of users interact in billions of combinations; the data sets
must be analyzed, summarized and presented visually.

@@

@@

\begin{tabular}[t]{lp{100mm}}
\textbf{JSON} &
JavaScript Object Notation, a lightweight text-based human-readable data
interchange format. It is used for representing simple data structures and
associative arrays (called objects). JSON is often used for serializing
structured data and transmitting it over network connection.

@@

\textbf{Replication} &
.

@@

\textbf{Sharding} &
.
\end{tabular}

\chapter{Text with no home and no current purpose}

\section{It's not like the movies}

Luckily, this is a problem that has already been solved. Google has published
papers on the MapReduce framework \cite{mapreduce}, which allows for parallel
processing of large data sets, and The Google File System \cite{gfs}, a
scalable, distributed file system. The Apache Hadoop \cite{hadoop} project has
published free software implementations of a MapReduce framework and a
distributed file system, both of which build on the ideas in the papers
published by Google.

\section{Techniques and tools}

TODO: Why do we chose HBase in favour of other techniques (i.e. SQL)?

In actual practice, the main difference between the two models is that
querying is highly restricted in a BigTable-like data store---there are no
joins, for example---whereas the relational database was born to be queried.
On the other hand, the restrictions placed upon the BigTable models are
what allows it to scale to such a large data set. Indeed, BigTable was born
to scale.

TODO: elaborate on reasons to why starsql does not scale

Many applications have the behaviour where they write data once and read it
a lot more, log analysers for instance. This writing-once-reading-many
behaviour is precisely what BigTable is designed for. BigTable is designed
for sequential, as opposed to random, data access. A typical use case is
the scan, where a large number of rows are read in sequence and a certain
column or column-family are picked up and returned. Certainly, you will
want to store those data points that are likely to be needed simultaneously
close to each other. \cite{bigtable}

Sparseness is a big thing. Null values are free of charge. TODO: *Describe
how we leverage sparsity*.

\section{Electromagnetic field theory}

-- A truly magnetic data store - the idea of enabling the user --

From a user's perspective, storing aggregates in Zohmg is as simple as writing
a map function (in the MapReduce sense) and letting Zohmg run that map function
over the data set. The user's map function is called \textit{user map function}
from here on. In order to let Zohmg know what output to expect from the mapper,
the user also defines the dimensions and units of his data. After all, the user
most likely knows his own data better than anyone else.

>>>>

\noindent {\Large \textbf{Abstract}}
\vspace{12pt}

@@

\noindent Analyzing data at a massive scale is one of the biggest challenges
that Last.fm is facing. Interpreting patterns in user behaviour becomes a
challenge when millions of users interact in billions of combinations; the data
sets must be analyzed, summarized and presented visually.

@@

\vspace{48pt}

\noindent {\Large \textbf{Sammanfattning}}
\vspace{12pt}

% TODO: sync with english abstract.

\noindent Att analysera data p{\aa} en massiv skala {\"a}r en av de st{\"o}rsta
utmaningarna som Last.fm st{\aa}r inf{\"o}r. Att tolka m{\"o}nster i
anv{\"a}ndarbeteende blir en utmaning n{\"a}r miljoner anv{\"a}ndare samspelar i
miljarder kombinationer. Datam{\"a}ngderna m{\aa}ste analyseras, summeras och
presenteras visuellt.

Detta examensarbete beskriver ett datalager f{\"o}r multidimensionell
tidsseriebaserad data. M{\aa}tt {\"a}r summerade {\"o}ver multipla dimensioner.
Datalagret {\"a}r optimerat f{\"o}r dataextraheringshastighet: Ett av
designm{\aa}len {\"a}r att servera data i musklickshastighet f{\"o}r att
fr{\"a}mja utforskning av data i realtid.

Liknande datalager existerar men de anv{\"a}nder oftast relationella
databassystem som databas f{\"o}r back-end. Nyheten i v{\aa}rt angripss{\"a}tt
{\"a}r att modellera multidimensionella datakuber ovanp{\aa} en distribuerad,
kolumnorienterad databas f{\"o}r att utnyttja skalbarhetsf{\"o}rdelarna av
s{\aa}dana databaser.

\pagebreak
\phantom{ }
\pagebreak

\noindent {\Large \textbf{Preface}}
\vspace{12pt}

\noindent Per Andersson and Fredrik M{\"o}llerstrand both study Computer Science
at Chalmers University of Technology. This report is part of their master thesis
degree. The work was performed at Last.fm's office in London during spring 2009.

The authors would like to specially thank the supervisors---David Sands at
Chalmers and Johan Oskarsson and Martin Dittus at Last.fm---for their help and
support.

The authors would also like to thank the staff at Last.fm and their family and
friends for moral support before, during, and after the thesis work.

\pagebreak
\phantom{ }
\pagebreak

@@

\begin{tabular}[t]{lp{100mm}}
\textbf{Column-oriented} &
A column-oriented database stores its content by column rather than by row.
This is an advantage for databases such as data warehouses, where aggregates
computed over large numbers of similar data items.

\textbf{JSON} &
JavaScript Object Notation, a lightweight text-based human-readable data
interchange format. It is used for representing simple data structures and
associative arrays (called objects). JSON is often used for serializing
structured data and transmitting it over network connection.

@@

\textbf{Replication} &
The process of sharing information for ensurance of consistency between
redundant sources. For instance multiple databases on different machines share
the same content for sake of redundancy, the information is not lost if one
machine fails.

@@

\textbf{Sharding} &
A method for horizontal partitioning in a database. Horizontal partinioning is a
design principle whereby rows of a database table are held seperately. Each
partition forms a so called \textit{shard}, which may be located on a separate
database server. The advantage is the number of rows in each table is reduced,
which reduces index size and thus performance.
\end{tabular}
